---
title: "Catalonia as an Electoral Weapon: Discursive Framing and Territorial Instrumentalization in the Parliament of Andalusia"
shorttitle: "Catalonia as an Electoral Weapon"
author:
  - name: Claudi Sonet
    affiliations:
      - name: Universitat d'Anthròpica
        department: Departament d'Antropomorfisme Aplicat
        address: San Francisco, CA, USA
  - name: Gemma Bessons
    affiliations:
      - name: Googleplex State
        department: Institut de Recerca en Constel·lacions de Dades
        address: Mountain View, CA, USA
  - name: Geppetto Openini
    affiliations:
      - name: Acadèmia de l'Oberta Intel·ligència
        department: Laboratori de Titelles Autònoms i Sintaxi de Fusta
        address: Altitut de l'Estel, USA
    email: pinocho.realboy@openai.lol
    corresponding: true
date: today
abstract: |
  Why does a regional parliament devote nearly one in twelve plenary interventions
  to another region's politics? This paper examines the discursive framing of
  Catalonia in the Parliament of Andalusia during the Twelfth Legislature
  (2022--2025), drawing on a corpus of 9,942 plenary interventions
  processed through an automated pipeline combining regex detection and large
  language model (LLM) classification. Following the measurement validation
  framework of @halterman_keith2025, we deploy a blind-coding protocol in which
  GPT classifies each intervention---without knowledge of the speaker's identity
  or party---along two dimensions: sentiment toward Catalonia (Hostile, Neutral,
  Empathetic) and primary discursive frame (Electoral Weaponization, Constitutional
  Threat, Grievance Economics, Positive Reference, Other). Of 792 detected mentions,
  62.6\% are coded as hostile, and 44.6\% as Electoral Weaponization. The data
  reveal a striking asymmetry between the right-wing bloc (PP: 78.6\% hostile;
  Vox: 85.7\% hostile) and left-wing parties (PSOE: 6.1\% hostile; Por Andalucía:
  9.4\% hostile). Within the right, we identify a discursive division of labour:
  the PP pivots toward Grievance Economics (25.9\%), linking Catalonia to regional
  financing grievances, while Vox concentrates on Constitutional Threat (31.2\%),
  mobilizing an alarmist discourse on national unity. We conceptualize this pattern
  as *strategic Catalanophobia*---the systematic deployment of a neighbouring
  region's political conflict as a substitute for domestic governance debate---and
  discuss its implications for the functioning of multilevel democracy in Spain.

keywords:
  - political framing
  - territorial politics
  - affective polarization
  - large language models
  - Spain
  - Andalusia
  - Catalonia

format:
  pdf:
    documentclass: article
    geometry: margin=1in
    fontsize: 12pt
    linestretch: 1.5
    number-sections: true
    toc: false
    toc-depth: 2
    colorlinks: true
    cite-method: biblatex
    biblio-style: authoryear
    keep-tex: false
    include-in-header:
      text: |
        \usepackage{booktabs}
        \usepackage{dcolumn}
        \usepackage{rotating}
        \usepackage{float}
        \floatplacement{figure}{H}

bibliography: references.bib
---

```{r setup}
#| include: false
#| cache: false

PROJ_ROOT <- normalizePath("../..")
knitr::opts_knit$set(root.dir = PROJ_ROOT)
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.pos = "H",
  out.width = "\\textwidth"
)

library(tidyverse)
library(lubridate)
library(scales)
library(knitr)
library(kableExtra)
library(patchwork)

# ── Load data ─────────────────────────────────────────────────────────────────
corpus <- read.csv(
  file.path(PROJ_ROOT, "02_dades_processades/dataset_consolidat/corpus_intervencions.csv"),
  stringsAsFactors = FALSE
) |>
  mutate(
    fecha              = as.Date(fecha),
    data_mes           = as.Date(data_mes),
    mencio_cat_ext     = as.logical(mencio_cat_ext),
    mencio_catalunya   = as.logical(mencio_catalunya),
    orador_identificat = as.logical(orador_identificat),
    data_trim          = floor_date(fecha, "quarter")
  )

progress <- read.csv(
  file.path(PROJ_ROOT, "02_dades_processades/llm_frames/llm_progress.csv"),
  stringsAsFactors = FALSE
) |>
  mutate(
    id_master         = suppressWarnings(as.integer(id_master)),
    intervencio_ordre = suppressWarnings(as.integer(intervencio_ordre))
  ) |>
  filter(!is.na(id_master), !is.na(intervencio_ordre)) |>
  select(-any_of(c("row_key", "llm_ts")))

llm <- corpus |>
  left_join(progress, by = c("id_master", "intervencio_ordre")) |>
  filter(!is.na(llm_sentiment_cat))

# ── Colour palettes ───────────────────────────────────────────────────────────
COL_SENTIMENT <- c(
  "Hostil_Negatiu"        = "#C0392B",
  "Neutral_Institucional" = "#95A5A6",
  "Empàtic_Positiu"       = "#27AE60"
)
COL_FRAME <- c(
  "Constitutional_Threat"   = "#7B241C",
  "Electoral_Weaponization" = "#6C3483",
  "Grievance_Economics"     = "#D35400",
  "Positive_Reference"      = "#1A8A6A",
  "Other"                   = "#BDC3C7"
)
FRAME_LVLS <- c("Electoral_Weaponization", "Constitutional_Threat",
                "Grievance_Economics", "Positive_Reference", "Other")

PARTY_ORDER <- c("Vox","PP","PSOE","Por Andalucía","Adelante Andalucía")

theme_paper <- theme_minimal(base_size = 10) +
  theme(
    plot.title    = element_text(face = "bold", size = 10),
    plot.subtitle = element_text(color = "grey40", size = 8),
    legend.position = "bottom",
    legend.key.size = unit(0.35, "cm"),
    legend.text = element_text(size = 7),
    axis.text   = element_text(size = 8),
    strip.text  = element_text(size = 8, face = "bold")
  )
theme_set(theme_paper)
```

\newpage
\tableofcontents

\newpage

# Introduction: The Andalusian Echo Chamber

Political communication in multilevel democracies is rarely confined to the
jurisdictional boundaries of a given legislature. Regional parliaments, while
institutionally mandated to address the specific concerns of their constituent
territories, frequently serve as arenas for national-level ideological conflict
[@astudillo_pallares2013]. Yet the extent to which this extraterritorial
displacement occurs---and the discursive mechanisms through which it operates---has
received comparatively little systematic attention in the scholarly literature.
This paper addresses that gap by examining a striking empirical puzzle: in the
Parliament of Andalusia, approximately 8 per cent of all plenary interventions
contain an explicit reference to Catalonia, rising to 15 per cent among Vox
parliamentarians.

The puzzle is analytically significant for at least three reasons. First,
Andalusia and Catalonia are geographically distant, economically distinct, and
institutionally autonomous communities with divergent political traditions.
There is no obvious *functional* reason why the Andalusian assembly should
devote substantial debating time to Catalan politics. Second, the volume of
references is not evenly distributed across the political spectrum: it is
overwhelmingly concentrated in the right-wing parties (PP and Vox), and
overwhelmingly hostile in tone. Third, the Twelfth Legislature (2022--present)
coincides with a period of particular salience: the PP holds an absolute majority
in Andalusia while the central government in Madrid depends on the parliamentary
support of Catalan independence parties, creating an acute structural tension
between national and subnational political logics.

We argue that these references do not represent spontaneous political commentary
but rather a coherent discursive strategy: the systematic instrumentalization of
Catalonia as a foil to frame domestic political competition, mobilize partisan
loyalties, and delegitimize rivals---a phenomenon we conceptualize as *strategic
Catalanophobia*. Drawing on a fully automated, methodologically validated pipeline
for corpus construction and LLM-based content classification, we provide the first
systematic empirical mapping of this phenomenon in the Twelfth Andalusian Legislature
(2022--2025).

The remainder of the paper is structured as follows. Section 2 develops the
theoretical framework, integrating framing theory, territorial grievance
literature, and the emerging scholarship on affective polarization in Spain.
Section 3 describes the data and methodology, including the measurement
validation protocol. Section 4 presents the empirical results. Section 5
offers a theoretical discussion and conclusions.

# Theoretical Framework

## Framing and Political Elites

The concept of framing in political communication refers to the process by which
political actors "select some aspects of a perceived reality and make them more
salient in a communicating text, in such a way as to promote a particular problem
definition, causal interpretation, moral evaluation, and/or treatment recommendation"
[@entman1993, p. 52]. Framing is not merely a matter of rhetoric: by activating
different cognitive schemas, elite frames shape how mass audiences interpret
political events and distribute attributions of responsibility [@chong_druckman2007].
The choice of frame is therefore inherently strategic, reflecting not only
ideological commitments but also calculations about electoral mobilization and
coalition maintenance.

In the context of multilevel polities, frame selection by regional legislators
takes on additional dimensions. Regional parliaments occupy an intermediate
position in the political system: their members are simultaneously local
representatives and members of national party organizations, making them
susceptible to pressures from both levels of the political hierarchy
[@astudillo_pallares2013]. When national-level partisan conflicts become
sufficiently salient, regional parliaments may serve as "resonance chambers"
that amplify and locally adapt central political narratives. The Andalusian
case offers an unusually clear instance of this dynamic, given the structural
asymmetry between the regional government (PP absolute majority) and the
national government (dependent on Catalan independence parties).

## Territorial Grievance and Interregional Competition

Spain's system of fiscal decentralization has historically generated significant
interregional distributional conflicts [@lopez_laborda2006]. The negotiation of
regional financing arrangements is a perennial source of political contention,
and references to Catalonia frequently appear in this context: the perception
that Madrid has granted fiscal concessions to Catalonia at the expense of other
communities---particularly Andalusia, Spain's most populous region---is a
recurring motif in Andalusian political discourse [@amat2012].

The grievance frame (*Grievance_Economics* in our coding scheme) mobilizes a
specific causal story: resources that should flow to Andalusia are being diverted
to Catalonia as the price of parliamentary support. This narrative has a
particularly strong popular resonance in a region with persistent structural
unemployment and high dependence on central government transfers. It also serves
a dual function: simultaneously legitimizing claims for greater regional resources
and delegitimizing the national government that is seen to be complicit in
Andalusia's disadvantage.

## Affective Polarization and the Territorial Cleavage

Recent scholarship on affective polarization in Spain has highlighted the extent
to which the territorial dimension---particularly the Catalan independence
conflict---has restructured the emotional landscape of Spanish politics
[@miller_torcal2020; @rodon2023]. The emergence of Vox as a major electoral force
since 2018 has intensified this dynamic: the party's foundational narrative is
organized around the defence of national unity against Catalan and Basque
secessionism, and its electoral success has pressured the PP to compete on
similar terrain [@meguid2005].

The concept of affective polarization [@iyengar_westwood2015] is directly
applicable to this context: Catalonia functions not merely as a policy debate
but as a negative reference point---an "out-group" whose presence in political
discourse serves primarily to consolidate in-group (right-wing, Spanish
nationalist) identities. On this reading, the frequency and hostility of
Catalan references in Andalusian plenary debates reflects not an intrinsic
interest in Catalan affairs but a strategic use of Catalonia to perform and
reinforce partisan belonging. The moral dimension of this conflict is further
analysed in @sanjaume2020, who situates the territorial dispute within broader
frameworks of political legitimacy and recognition.

# Data and Methodology

## Corpus Construction

Our corpus comprises all plenary interventions recorded in the videoteca
(video archive) of the Parliament of Andalusia for the Twelfth Legislature
(2022--2025), the current legislature in which the PP holds an absolute majority.
The data were obtained through a multi-stage automated pipeline. In the first
stage, a web scraper extracted the catalogue of plenary sessions (N = 189
sessions) and resolved the internal session identifiers (`id_master`) by
visiting each video page and parsing a JSPLAYLIST block embedded in the HTML
source. In the second stage, a session-level API was queried to obtain the
hierarchical structure of each session (agenda points and speakers), and a
speaker-level API was queried to obtain full speech-to-text (STT) transcriptions
for each individual intervention. The resulting corpus contains **9,942 plenary
interventions** spanning three years of parliamentary activity.

Each intervention was matched, via normalized name-matching, to a registry of
deputies and government members, enabling the assignment of party affiliation,
parliamentary group, and legislature. Government members (consejeros and consejeras
of the Junta de Andalucía) who appeared in plenary sessions without holding
a parliamentary seat were identified separately and assigned to their respective
parties.

## Detection of Catalan References

Initial detection of interventions containing references to Catalonia was
performed using a regular expression (REGEX) designed to capture both direct
toponym references and associated political vocabulary:

$$\texttt{(?i)(catalu[\~{n}n]a \,|\, catalan(es|as|\'{a}n)? \,|\, independentis(mo|tas?) \,|\, separatis(mo|tas?) \,|\, golpistas?)}$$

This extended pattern was developed inductively: a preliminary regex
(`catalu[ñn]a`) was applied to extract a sub-corpus of 200-word context
windows around each initial match, which was then examined to identify
additional lexical forms used to refer to Catalonia or the independence
movement. The final regex detects 792 interventions (8.0\% of the corpus),
ranging from 15.0\% of Vox interventions to 6.0\% of PSOE interventions
(see Figure 1).

## LLM Classification and Measurement Validation

The 792 detected interventions were classified using GPT (OpenAI), following
the measurement validation framework proposed by @halterman_keith2025 for
the use of LLMs as coding instruments in political science. Two core
methodological safeguards were implemented.

First, **blind coding**: the model received only the text of each
intervention, with no information about the speaker's identity, party
affiliation, or legislative context. This design eliminates the risk that
the model's output reflects prior knowledge about the speaker rather than
the textual content of the intervention---a form of construct invalidity
that has been identified as a significant concern in LLM-based annotation
[@grimmer_stewart2013].

Second, **structured output**: the model was constrained to return a JSON
object conforming to a predefined schema, enforcing the valid value sets for
each variable. This eliminates free-text hallucination and ensures
categorical consistency across all 792 classifications.

The classification instrument (codebook) covers two dimensions:

**Sentiment** (`sentiment_cat`): the affective orientation of the intervention
toward Catalan institutions, the Catalan political movement, or Catalan society.
Three categories are distinguished: *Hostil_Negatiu* (hostile/negative:
pejorative, alarmist, or delegitimizing language); *Neutral_Institucional*
(neutral: factual references, statistics, procedural discussion); and
*Empàtic_Positiu* (empathetic/positive: solidarity, Catalonia as a positive
reference, dialogue-oriented language).

**Primary Frame** (`primary_frame`): the dominant discursive lens through which
the speaker introduces Catalonia into the Andalusian debate. Five categories
are distinguished: *Electoral_Weaponization* (Catalan pacts used to attack
national rivals); *Constitutional_Threat* (Catalonia as a threat to national
unity or the rule of law); *Grievance_Economics* (Andalusia as fiscally
disadvantaged by concessions to Catalonia); *Positive_Reference* (Catalonia
as a positive policy model); and *Other* (references that fit none of the
above categories).

Each classification also includes a *chain of thought* field in which the
model articulates its reasoning with explicit reference to the codebook's
exclusion rules, and a *confidence score* (0--1). The mean confidence across
all 792 classifications is 0.88 (SD = 0.08), suggesting high model certainty.

# Results: The Anatomy of Hostility

## The Intensity and Distribution of Catalan References

Figure 1 displays the proportion of interventions containing Catalan references
by party. The variation across the political spectrum is striking. Vox records
the highest rate (15.0\%, N = 112), followed by Por Andalucía (9.8\%, N = 64)
and Adelante Andalucía (8.1\%, N = 30). The PP, as the dominant party in the
legislature, records 7.6\% (N = 490)---a lower rate than Vox but a substantially
larger absolute volume. The PSOE, in opposition, records the lowest rate (6.0\%,
N = 82). The temporal distribution (Figure 2) reveals that references are not
uniformly distributed across the legislature but cluster around specific political
events: the approval of national budgets, debates on regional financing, and episodes
of political tension between the central government and Catalan parties.

```{r fig-pct-partit}
#| fig-cap: "Proportion of interventions containing Catalan references, by party. The horizontal axis shows the percentage of each party's total plenary interventions that contain at least one match to the extended Catalonia regex. N denotes the absolute number of interventions classified."
#| fig-height: 3.2

corpus |>
  filter(!is.na(partit)) |>
  group_by(partit) |>
  summarise(n = n(), n_cat = sum(mencio_cat_ext, na.rm = TRUE),
            pct = 100*n_cat/n, .groups = "drop") |>
  mutate(
    partit = factor(partit, levels = rev(PARTY_ORDER)),
    label  = sprintf("%.1f%% (N=%d)", pct, n_cat)
  ) |>
  ggplot(aes(pct, partit)) +
  geom_col(fill = "#2C3E50", alpha = 0.8, width = 0.6) +
  geom_text(aes(label = label), hjust = -0.08, size = 2.8) +
  scale_x_continuous(labels = label_percent(scale = 1),
                     expand = expansion(mult = c(0, 0.22)),
                     limits = c(0, 21)) +
  labs(x = "% of party interventions", y = NULL)
```

```{r fig-temporal}
#| fig-cap: "Temporal evolution of Catalan references. Upper panel: absolute number of interventions per month. Lower panel: proportion of monthly interventions containing Catalan references, with a LOESS smoother (shaded 95\\% confidence interval)."
#| fig-height: 4.2

cat_mes <- corpus |>
  filter(!is.na(data_mes)) |>
  group_by(data_mes) |>
  summarise(n_total = n(), n_cat = sum(mencio_cat_ext, na.rm = TRUE),
            pct = 100*n_cat/n_total, .groups = "drop")

p1 <- ggplot(cat_mes, aes(data_mes, n_cat)) +
  geom_col(fill = "#2C3E50", alpha = 0.75, width = 20) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(y = "N interventions", x = NULL,
       title = "Monthly count of interventions with Catalan references")

p2 <- ggplot(cat_mes, aes(data_mes, pct)) +
  geom_line(color = "#2C3E50", linewidth = 0.8) +
  geom_smooth(method = "loess", se = TRUE, color = "#C0392B",
              fill = "#C0392B", alpha = 0.15, linewidth = 0.6) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(labels = label_percent(scale = 1), limits = c(0, NA)) +
  labs(y = "% of monthly interventions", x = NULL,
       title = "Proportion of interventions with Catalan references")

p1 / p2
```

## The Sentiment Gap: Hostility as a Right-Wing Characteristic

The most striking finding of the analysis concerns the distribution of sentiment
across the political spectrum. At the aggregate level, 62.6\% of all classified
interventions are coded as *Hostil_Negatiu*, 32.7\% as *Neutral_Institucional*,
and a mere 4.7\% as *Empàtic_Positiu*. However, these global figures conceal a
profound asymmetry between the right and left blocs.

Table 1 and Figure 3 present the sentiment distribution by party. Among right-wing
parties, hostility is overwhelmingly dominant: Vox records 85.7\% hostile
interventions and the PP records 78.6\%. By contrast, the left-wing parties
exhibit markedly different profiles: the PSOE records only 6.1\% hostile
interventions (with 68.3\% neutral and 25.6\% empathetic), while Por Andalucía
records 9.4\% hostile (76.6\% neutral, 14.1\% empathetic) and Adelante Andalucía
6.7\% hostile (83.3\% neutral, 10.0\% empathetic). The gap between the most
hostile party (Vox, 85.7\%) and the least hostile (PSOE, 6.1\%) is 79.6
percentage points---a difference of extraordinary magnitude that suggests the
two blocs are operating in almost entirely different discursive registers when
referring to Catalonia.

```{r tbl-sentiment}
#| tbl-cap: "Sentiment distribution by party (percentage of classified interventions). N denotes the total number of LLM-classified interventions per party. Rows are sorted by descending percentage of hostile interventions."

llm |>
  filter(!is.na(partit)) |>
  group_by(Partit = factor(partit, levels = PARTY_ORDER)) |>
  summarise(
    N           = n(),
    `Hostile (%)` = round(100*mean(llm_sentiment_cat=="Hostil_Negatiu"),   1),
    `Neutral (%)` = round(100*mean(llm_sentiment_cat=="Neutral_Institucional"),1),
    `Empathetic (%)` = round(100*mean(llm_sentiment_cat=="Empàtic_Positiu"),1),
    .groups = "drop"
  ) |>
  arrange(desc(`Hostile (%)`)) |>
  kable(booktabs = TRUE, linesep = "",
        align = c("l","r","r","r","r")) |>
  kable_styling(latex_options = c("hold_position","striped"),
                font_size = 9, full_width = FALSE)
```

```{r fig-sentiment}
#| fig-cap: "Sentiment distribution by party (stacked bar, percentage of classified interventions). Parties are ordered by descending proportion of hostile interventions."
#| fig-height: 3.2

llm |>
  filter(!is.na(partit)) |>
  count(partit, llm_sentiment_cat) |>
  group_by(partit) |>
  mutate(pct = 100*n/sum(n)) |>
  ungroup() |>
  mutate(
    partit = factor(partit, levels = rev(PARTY_ORDER)),
    llm_sentiment_cat = factor(llm_sentiment_cat,
      levels = c("Hostil_Negatiu","Neutral_Institucional","Empàtic_Positiu"),
      labels = c("Hostile","Neutral","Empathetic"))
  ) |>
  ggplot(aes(pct, partit, fill = llm_sentiment_cat)) +
  geom_col(position = "stack", alpha = 0.9, width = 0.65) +
  geom_text(aes(label = if_else(pct > 5, paste0(round(pct), "%"), "")),
            position = position_stack(vjust = 0.5),
            size = 2.6, color = "white", fontface = "bold") +
  scale_fill_manual(values = c("Hostile" = "#C0392B",
                               "Neutral" = "#95A5A6",
                               "Empathetic" = "#27AE60"),
                    name = "Sentiment") +
  scale_x_continuous(labels = label_percent(scale = 1),
                     expand = expansion(mult = c(0, 0.02))) +
  labs(x = "% of classified interventions", y = NULL)
```

## Divergence within the Right: A Discursive Division of Labour

Among parties on the right, the sentiment picture is one of shared hostility; yet
the *type* of hostility differs substantially. This divergence is captured by the
primary frame variable.

Figure 4 and Table 2 reveal that both PP and Vox share *Electoral Weaponization*
as their most frequent frame (50.0\% each), using Catalan political dynamics
primarily to attack the national government and its left-wing supporters. However,
their secondary frames diverge sharply: the PP concentrates 25.9\% of its
interventions in *Grievance_Economics*---a frame that connects Catalonia to the
regional financing debate and Andalusia's fiscal disadvantage---while Vox
concentrates 31.2\% in *Constitutional_Threat*, a frame that foregrounds alarm
about national unity and the legitimacy of the political order.

This divergence is substantively significant. The PP's *Grievance_Economics*
frame is essentially redistributive: it mobilizes Andalusian identity against
perceived fiscal injustice, positioning the PP as the defender of Andalusian
interests against a national government that sacrifices regional equity for
coalition survival. Vox's *Constitutional_Threat* frame is instead
identitarian-authoritarian: it mobilizes Spanish nationalist sentiment by
representing Catalonia as an existential menace to the constitutional order.
The two frames appeal to distinct emotional registers and activate different
aspects of the territorial cleavage---economic resentment in the PP's case,
and national security anxiety in Vox's.

For the left-wing parties, the frame distribution is also distinctive: the
PSOE records 53.7\% in the *Other* category (reflecting mentions that are
largely incidental or procedural), and Por Andalucía similarly records 60.9\%
in *Other*. Electoral Weaponization is present on the left as well (PSOE:
29.3\%; Adelante Andalucía: 30.0\%), but it operates in the reverse direction:
these parties use Catalan references to attack the PP and Vox for their alleged
manipulation of the territorial issue for partisan ends.

```{r fig-frame}
#| fig-cap: "Primary discursive frame by party (stacked bar, percentage of classified interventions). Parties are ordered by descending proportion of adversarial frames (Electoral Weaponization + Constitutional Threat)."
#| fig-height: 3.5

llm |>
  filter(!is.na(partit)) |>
  count(partit, llm_primary_frame) |>
  group_by(partit) |>
  mutate(pct = 100*n/sum(n)) |>
  ungroup() |>
  mutate(
    adv = ifelse(llm_primary_frame %in%
                   c("Electoral_Weaponization","Constitutional_Threat"), pct, 0)
  ) |>
  group_by(partit) |>
  mutate(adv_total = sum(adv)) |>
  ungroup() |>
  mutate(
    partit = fct_reorder(factor(partit, levels = PARTY_ORDER), adv_total),
    llm_primary_frame = factor(llm_primary_frame, levels = rev(FRAME_LVLS),
      labels = rev(c("Electoral Weaponization","Constitutional Threat",
                     "Grievance Economics","Positive Reference","Other")))
  ) |>
  ggplot(aes(pct, partit, fill = llm_primary_frame)) +
  geom_col(position = "stack", alpha = 0.9, width = 0.65) +
  geom_text(aes(label = if_else(pct > 6, paste0(round(pct), "%"), "")),
            position = position_stack(vjust = 0.5),
            size = 2.5, color = "white", fontface = "bold") +
  scale_fill_manual(
    values = c(
      "Electoral Weaponization" = "#6C3483",
      "Constitutional Threat"   = "#7B241C",
      "Grievance Economics"     = "#D35400",
      "Positive Reference"      = "#1A8A6A",
      "Other"                   = "#BDC3C7"
    ), name = "Frame") +
  scale_x_continuous(labels = label_percent(scale = 1),
                     expand = expansion(mult = c(0, 0.02))) +
  labs(x = "% of classified interventions", y = NULL)
```

```{r tbl-frame}
#| tbl-cap: "Primary discursive frame by party (percentage of classified interventions). EW = Electoral Weaponization; CT = Constitutional Threat; GE = Grievance Economics; PR = Positive Reference. N denotes total classified interventions. Rows sorted by descending adversarial frame proportion."

llm |>
  filter(!is.na(partit)) |>
  count(Partit = factor(partit, levels = PARTY_ORDER), frame = llm_primary_frame) |>
  group_by(Partit) |>
  mutate(pct = round(100*n/sum(n), 1), N = sum(n)) |>
  ungroup() |>
  pivot_wider(id_cols = c(Partit, N), names_from = frame,
              values_from = pct, values_fill = 0) |>
  select(Partit, N,
         `EW (%)` = any_of("Electoral_Weaponization"),
         `CT (%)` = any_of("Constitutional_Threat"),
         `GE (%)` = any_of("Grievance_Economics"),
         `PR (%)` = any_of("Positive_Reference"),
         `Other (%)` = any_of("Other")) |>
  arrange(desc(rowSums(across(any_of(c("EW (%)","CT (%)")), \(x) replace_na(x,0))))) |>
  kable(booktabs = TRUE, linesep = "",
        align = c("l","r","r","r","r","r","r")) |>
  kable_styling(latex_options = c("hold_position","striped"),
                font_size = 9, full_width = FALSE)
```

## Temporal Dynamics: Intensification and Electoral Cycles

The temporal evolution of Catalan references (Figure 2) is not monotone. A
notable intensification is visible during 2023, coinciding with the national
electoral cycle (general elections of July 2023) and the subsequent negotiations
over the investiture of the central government---a process that involved
controversial agreements with Catalan independence parties, including a contested
amnesty law. This pattern is consistent with the Electoral Weaponization frame:
the salience of Catalan references in the Andalusian parliament tracks, to a
significant degree, the salience of Catalonia in national-level political conflict.

Figure 5 disaggregates the temporal evolution by party and sentiment. The PP
and Vox display consistently high rates of hostile framing throughout the
legislature, with notable spikes during the periods described above. Left-wing
parties display broadly stable, low rates of hostility, with some increase in
empathetic framing during periods of dialogue between the central government
and Catalan institutions.

```{r fig-temporal-sent}
#| fig-cap: "Quarterly evolution of hostile sentiment by party (percentage of classified interventions per quarter). Only quarters with at least three classified interventions are shown."
#| fig-height: 3.8

llm |>
  filter(!is.na(partit), !is.na(data_trim)) |>
  group_by(data_trim, partit = factor(partit, levels = PARTY_ORDER)) |>
  summarise(n_tot = n(),
            pct_host = 100*mean(llm_sentiment_cat == "Hostil_Negatiu"),
            .groups = "drop") |>
  filter(n_tot >= 3) |>
  ggplot(aes(data_trim, pct_host, color = partit)) +
  geom_line(linewidth = 0.7, alpha = 0.85) +
  geom_point(size = 1.5, alpha = 0.8) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(labels = label_percent(scale = 1), limits = c(0, 105)) +
  labs(x = NULL, y = "% hostile interventions",
       color = "Party",
       title = "Quarterly proportion of hostile sentiment toward Catalonia, by party")
```

# Discussion: Strategic Catalanophobia

## Conceptualizing the Pattern

The results presented above support a coherent interpretive framework that we
term *strategic Catalanophobia*. The concept denotes the systematic, strategically
motivated deployment of negative discursive representations of Catalonia---its
politics, institutions, and political movement---as a mechanism for achieving
domestic political goals in a regional parliament. The term "strategic" is
important: the hostility toward Catalonia visible in Andalusian plenary debates
is not a spontaneous or incidental expression of political sentiment but a
deliberate communicative choice, reflected in the consistency of the pattern
across speakers, parliamentary groups, and issue contexts.

Strategic Catalanophobia serves multiple, overlapping functions within the
Andalusian political field. First, it provides right-wing parties with a
*legitimizing narrative for governance*: by attributing Andalusia's fiscal
difficulties to Catalan exceptionalism and the national government's capitulation
to Catalan demands, the PP can mobilize a discourse of regional victimhood that
simultaneously positions Andalusia as disadvantaged and the PP as its defender.
Second, it serves as *opposition weaponization*: the PSOE, in government at the
national level, is held accountable in the Andalusian parliament for decisions
made in Madrid, creating a form of vertical displacement of political
accountability that complicates the attribution of responsibility in citizens'
minds. Third, for Vox, it provides a *nationalist mobilization frame*: by
representing Catalonia as an existential threat to Spain, Vox activates the
constitutional-emotional register of Spanish nationalism and distinguishes itself
from the PP through the radicalism of its framing.

## Implications for Multilevel Democracy

The phenomenon of strategic Catalanophobia raises broader questions about the
health of multilevel democracy in Spain. When regional parliaments function
primarily as stages for national-level partisan combat---importing the conflicts
of Madrid into the deliberations of Seville---they risk losing their distinctive
function as arenas for territorial self-governance [@astudillo_pallares2013].
The substitution of substantive debate about regional policy for performative
confrontation about national political conflicts represents a form of deliberative
impoverishment that may have consequences for the quality of regional democratic
representation.

Moreover, the consistent association of Catalonia with hostility in the dominant
discursive output of Andalusian parliamentarians contributes to a wider process
of affective polarization along territorial lines [@miller_torcal2020; @rodon2023].
Citizens who receive their political information through the filter of Andalusian
parliamentary debate are exposed to an overwhelmingly negative representation of
Catalan politics, which may reinforce negative inter-territorial attitudes and
reduce the political space available for constructive dialogue about Spain's
constitutional architecture [@sanjaume2020].

The divergence between the PP and Vox within the right-wing bloc---the PP's
emphasis on economic grievance and Vox's emphasis on constitutional threat---suggests
that competitive dynamics within the right may drive a process of discursive
radicalization, as parties seek to differentiate their Catalan-reference strategies
in response to electoral pressure [@meguid2005]. The PP's pivot toward the
Grievance Economics frame, in particular, may represent an attempt to maintain
credibility with an Andalusian electorate for whom fiscal fairness is a concrete,
material concern, while simultaneously appropriating a theme (Catalan privilege)
that is central to Vox's electoral appeal.

## Methodological Considerations

Our analysis relies on LLM-based classification of textual content, a methodology
that is rapidly gaining traction in political science but remains subject to
important caveats [@grimmer_stewart2013; @halterman_keith2025]. Several design
choices deserve explicit consideration. The blind-coding protocol---in which the
model receives no information about the speaker's party or identity---eliminates
a significant potential source of bias, but it also means that the model cannot
leverage contextual information that a human coder might use to interpret ambiguous
passages. The structured output constraint ensures categorical consistency but
may impose artificial discretization on phenomena that are inherently continuous.
And the use of a single-model classification system---rather than a multi-model
ensemble or a human-LLM hybrid---limits the ability to assess inter-rater
reliability in traditional terms.

The mean confidence score of 0.88 across all classifications suggests that the
vast majority of interventions present unambiguous coding decisions; nonetheless,
a systematic validation exercise---such as the comparison of LLM outputs with
expert human codings on a random sample---would be required before the findings
are treated as definitive. We regard the present analysis as a substantively
illuminating exploratory study whose conclusions warrant subsequent validation.

# Conclusion

This paper has provided the first systematic empirical analysis of how Catalonia
is discursively framed in the Parliament of Andalusia. Our core findings can
be summarized as follows. References to Catalonia constitute 8 per cent of all
plenary interventions and reach 15 per cent among Vox parliamentarians. Among
classified interventions, 62.6 per cent are hostile in sentiment and 44.6 per cent
deploy an Electoral Weaponization frame. The sentiment gap between the right
(78--86\% hostile) and left (6--9\% hostile) blocs is extraordinary in
magnitude. Within the right, PP and Vox share Electoral Weaponization as their
primary frame but diverge in secondary framing---the PP toward Grievance
Economics, Vox toward Constitutional Threat---suggesting a discursive division
of labour that reflects their distinct electoral positions and base constituencies.

We have conceptualized these patterns under the rubric of *strategic Catalanophobia*:
a deliberate communicative strategy that serves simultaneously to mobilize
regional identity, attack national rivals, and avoid substantive debate about
domestic governance. This strategy is not unique to Andalusia, but the
combination of Andalusia's size (Spain's most populous autonomous community),
the PP's absolute majority, and the structural tension between regional and
national government makes the Andalusian parliament an unusually clear and
consequential instance of this phenomenon.

Future research should extend the analysis to other regional parliaments, enabling
comparative assessment of the extent to which strategic Catalanophobia is a
specifically Andalusian phenomenon or a general feature of right-wing regional
politics in Spain. The methodology developed here---combining STT transcription,
regex-based corpus construction, and validated LLM classification---offers a
replicable template for this comparative enterprise.

\newpage

# References

::: {#refs}
:::
